---
title: "Week Sequence Analysis Figures"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Load packages and source file

```{r}
library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
library(cowplot)
library(tidyr)
source('src/plot_functions.R')
source('scripts/run_mat_hier.R')
```

Note: to run this, you can either directly load in nTreeClus results files in results/amdc_week_ntrees.csv OR recreate this saved file by running nTreeClus on daynamica data with run_nTreeClus_daynamica.ipynb.

## Run/load in clustering results and prep data

```{r}
out2 <- run_sim('clean_data/clean_week_seqs.csv', 'clean_data/clean_week_distmat.rds', 2, 9)
metric2 <- out2[[1]]
df2 <- out2[[2]]
ntrees_df2 <- read_csv('results/amdc_week_ntrees.csv') %>%
  rename(seq = seqs)
```

```{r}
df2 <- df2 %>%
  rename(seq = seqs) %>%
  mutate(seq_num = seq(1, nrow(df2))) %>%
  mutate(uniq_id = paste0(user_id, '_', seq_num))
ntrees_df2 <- ntrees_df2 %>%
  rename(seq_num = `...1`) %>%
  mutate(uniq_id = paste0(user_id, '_', seq_num))
hier_las2 <- create_lasagna_df('clust4', df2)
mat_las2 <- create_lasagna_df('clust6_EV12', df2)
ntrees_las2 <- create_lasagna_df('ntrees_DT', ntrees_df2)
```

### Get legend + white space for all plots

```{r}
plot_df <- mat_las2 %>%
  mutate(State = ifelse(State == 'H', 'Home', ifelse(State == 'O', 'Out',
        ifelse(State == 'T', 'Transport', ifelse(State == 'W', 'Work', NA)))))
p1 <- plot_lasagna(plot_df, 'clust6_EV12', 1) + theme(legend.position="bottom", legend.title=element_text(size=15), 
    legend.text=element_text(size=14))
legend_b <- cowplot::get_plot_component(p1, 'guide-box-bottom')

white_space <- ggplot() +
  theme_void()
```

## Contribution of each adjacency matrix compartment to the principal axes

```{r}
week_df <- read_csv('clean_data/clean_week_seqs.csv', show_col_types = FALSE)
indices <- list(c(1))
type <- c('keep')
trans_res <- final_gen(week_df$seqs, indices, type)

trans_mat <- trans_res[["vecs"]]$`1`
trans_mat_s <- scale(trans_mat, scale = FALSE)
svd_res <- svd(trans_mat_s)

vec <- 100/svd_res$d^2
XV2 <- (trans_mat_s %*% svd_res$v)^2
indv_contrib <- t(t(XV2)*vec)
rownames(indv_contrib) <- apply(expand.grid(c('H', 'O', 'T', 'W'), c('H', 'O', 'T', 'W')), 1, function(x){paste0(x, collapse = '')})
indv_contrib[,1:2]
```

## Figure 3

```{r, warning=FALSE, fig.height = 5, fig.width =7}
mat1 <- plot_lasagna_week(mat_las2, 'clust6_EV12', 3) +
  theme(legend.position="none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
mat2 <- plot_lasagna_week(mat_las2, 'clust6_EV12', 4) +
  theme(legend.position="none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
mat3 <- plot_lasagna_week(mat_las2, 'clust6_EV12', 2) +
  theme(legend.position="none", axis.text.x = element_blank(), axis.ticks.x = element_blank())

mat_p <- plot_grid(mat1, mat2, mat3, labels = c('A) 42.1%', 'B) 18.2%', 'C) 11.1%'), nrow = 2,label_y = 1.04, label_x = -0.01, label_size = 11)

final <- plot_grid(white_space, mat_p, legend_b, white_space,  axis = 'l', nrow = 4, rel_heights = c(1, 20, 1,1))
final
```

## Supplementary Figure 8

```{r, warning=FALSE, fig.width = 10, fig.height = 5}
sort_vals <- ntrees_df2 %>%
  group_by(ntrees_DT)%>%
  summarize(n = n()/nrow(df2)) %>%
  arrange(desc(n))

all_plots <- lapply(sort_vals$ntrees_DT, function(x){
  plot <- plot_lasagna_week(ntrees_las2, 'ntrees_DT', x) +
  theme(legend.position="none",  axis.text.x = element_blank(), axis.ticks.x = element_blank())
})

percs <- round(sort_vals$n, 3)*100
letters <- c('A) ', 'B) ', 'C) ' )
labels <- vector(length = length(percs))
for(i in seq(1,length(percs))){labels[i] <- paste0(letters[i], as.character(percs[i]), '%')}

mat_p <- plot_grid(plotlist = all_plots, labels = labels, nrow =1,label_y = 1.04, label_x = -0.01, label_size = 13)

final <- plot_grid(white_space, mat_p, white_space, legend_b, white_space,  axis = 'l', nrow = 5, rel_heights = c(0.75, 7, 0.75, .4,0.8))
final
```

## Supplementary Figure 9

```{r, warnings = FALSE, fig.height=12, fig.width = 8}
sort_vals <- df2 %>%
  group_by(clust6_EV12)%>%
  summarize(n = n()/nrow(df2)) %>%
  arrange(desc(n))

all_plots <- lapply(sort_vals$clust6_EV12, function(x){
  plot <- plot_lasagna_week(mat_las2, 'clust6_EV12', x) +
  theme(legend.position="none",  axis.text.x = element_blank(), axis.ticks.x = element_blank())
})

percs <- round(sort_vals$n, 3)*100
percs[2] <- 18.2 #so it sums to 100
letters <- c('A) ', 'B) ', 'C) ' , 'D) ', 'E) ', 'F) ')
labels <- vector(length = length(percs))
for(i in seq(1,length(percs))){labels[i] <- paste0(letters[i], as.character(percs[i]), '%')}

mat_p <- plot_grid(plotlist = all_plots, labels = labels, nrow = 3,label_y = 1.04, label_x = -0.01, label_size = 13)

final <- plot_grid(white_space, mat_p, white_space, legend_b, white_space,  axis = 'l', nrow = 5, rel_heights = c(0.75, 20, 0.75, .4,0.8))
final
```
